Spin canonical rings of log stacky curves

### 2015-07-10 surfaces-write-up/surface-write-up.tex
added relations gernal lemmas

\newcommand\nc{\newcommand}
\nc\on{\operatorname}
\nc\renc{\renewcommand}
\newcommand\ssec{\subsection}
\newcommand\sssec{\subsubsection}
\newcommand\bh{{\mathbb H}}
\newcommand\bn{{\mathbb N}}
\newcommand\bc{{\mathbb C}}
\newcommand\bbf{{\mathbb F}}
\newcommand\br{{\mathbb R}}
\newcommand\bq{{\mathbb Q}}
\newcommand\bp{{\mathbb P}}
\newcommand\bz{{\mathbb Z}}
\newcommand\ba{{\mathbb A}}

@@

\begin{prop}
\label{prop:cone-generation}
Let $n,t \in \bz,$ let $\alpha_1, \ldots, \alpha_n \in \bq$ and let $k_i^j \in \bz$ with $1 \leq i \leq n, 1 \leq j \leq t.$ Define
\begin{align*}
	\Sigma = \{(d,c_1, \ldots, c_n) \in \bz^{n+1} : c_i \geq - d \alpha_i,1 \leq i \leq n \text{ and } \sum_{i=1}^{n}k_i^j = 0, 1 \leq j \leq t\}
\end{align*}
Let $e_1, \ldots, e_r \in \Sigma,$ with $e_i = (\delta^i, c_1^i, \ldots, c_n^i)$ be a set of {\bf extremal rays} of $\Sigma$, meaning that the $\Sigma$ is contained in the $R_{\geq 0}$ span of $e_1, \ldots, e_n$.  Then, as a semigroup, $\Sigma$ is generated by elements whose first coordinate is less than $\sum_{i=1}^{n}\delta_i$.
\end{prop}
\begin{proof}
By assumption, $\sigma \in \Sigma$ as $\sigma = \sum_{i=1}^{n}a_i e_i$ with $a_i \in \br$. Let $\{a\} := a - \lfloor a \rfloor$ denote the fractional part of $a$. Let $\lambda = \sum_{i=1}^{n}\{a_i\}$. Hence, we can write $\sigma = \lambda + \sum_{i=1}^{n}\lfloor a_i \rfloor e_i.$ Hence, $\sigma$ lies in the $\bz_{\geq 0}$ span of $\lambda, e_1,\ldots, e_n$, which all have first coordinate less than $\sum_{i=1}^{n}\delta_i$. Therefore, $\Sigma$ is generated by elements whose first coordinate is less than $\sum_{i=1}^{n}\delta_i$.
\end{proof}

@@

\begin{prop}
\label{prop:pm-span-and-basis}
The functions $u^d \cdot \prod_{i=1}^n f_i^{c_i}$ so that 
\begin{align}
\label{align:pm-span}
\sum_{i=0}^{n} c_i \cdot d_i \cdot f_i = 0 & \text{ and } &c_i \geq \lfloor \alpha_i d\rfloor	
\end{align}
form a spanning set for $H^0(\bp^m, dD)$ over $k$. Furthermore, functions 
of the form $u^d \cdot \prod_{i=1}^n f_i^{c_i} $ such that
\begin{align}
\label{align:pm-basis}
c_i = -\lfloor d\alpha_i \rfloor & \text{ for } & i > m
\end{align}
form a basis for $H^0(\bp^m, dD)$ over $k$.
\end{prop}
\begin{proof}
By definition of $H^0(\bp^m,dD)$, functions satisfying conditions 
~\eqref{align:pm-span} lie in $H^0(\bp^m,dD)$. To complete the proof, it suffices to check functions satisfying conditions ~\eqref{align:pm-basis} form a basis of $H^0(\bp^m,dD)$. Note that there $\binom{m+ \lfloor dD \rfloor }{m}$ functions satisfying condition ~\eqref{align:pm-basis}. However we know $h^0(\bp^m,dD) = \binom{m+ \lfloor dD \rfloor }{m},$ so it suffices to show that those functions satisfying condition ~\eqref{align:pm-basis} are independent. This follows from the assumption that $f_0,\ldots, f_n$ form a basis of degree 1 rational functions, and so degree $\lfloor d \deg D \rfloor $ monomials in $f_0,\ldots, f_n$ form a basis of degree $\lfloor d \deg D \rfloor $ rational functions.
\end{proof}

\begin{rem}
Remark about how we can assume divisor only consists of hyperplanes
by taking the Veronese embedding.
\todo{do this}
\end{rem}

@@

>>>>

\newcommand\nc{\newcommand}
\nc\on{\operatorname}
\nc\renc{\renewcommand}
\newcommand\ssec{\subsection}
\newcommand\sssec{\subsubsection}
\newcommand\bh{{\mathbb H}}
\newcommand\bn{{\mathbb N}}
\newcommand\bc{{\mathbb C}}
\newcommand\bbf{{\mathbb F}}
\newcommand\br{{\mathbb R}}
\newcommand\bq{{\mathbb Q}}
\newcommand\bp{{\mathbb P}}
\newcommand\bz{{\mathbb Z}}
\newcommand\ba{{\mathbb A}}
\newcommand\bk{{\Bbbk}}

@@

\begin{prop}
\label{prop:cone-generation}
Let $n,t \in \bz,$ let $\alpha_1, \ldots, \alpha_n \in \bq$ and let $k_i^j \in \bz$ with $1 \leq i \leq n, 1 \leq j \leq t.$ Define
\begin{align*}
	\Sigma = \{(d,c_1, \ldots, c_n) \in \bz^{n+1} : c_i \geq - d \alpha_i,1 \leq i \leq n \text{ and } \sum_{i=1}^{n}k_i^j = 0, 1 \leq j \leq t\}
\end{align*}
Let $e_1, \ldots, e_r \in \Sigma,$ with $e_i = (\delta^i, c_1^i, \ldots, c_n^i)$ be a set of {\bf extremal rays} of $\Sigma$, meaning that the $\Sigma$ is contained in the $R_{\geq 0}$ span of $e_1, \ldots, e_r$.  Then, as a semigroup, $\Sigma$ is generated by elements whose first coordinate is less than $\sum_{i=1}^{n}\delta_i$. Furthermore, every element $\sigma \in \Sigma$ can be written in a canonical form $\lambda + \sum_{i=1}^{r} w_i e_i,$ with the first coordinate of $\lambda$ less than $\sum_{i=1}^{r}\delta_i$ and $w_1,\ldots, w_r \in \bz$.
\end{prop}
\begin{proof}
By assumption, $\sigma \in \Sigma$ as $\sigma = \sum_{i=1}^{r}a_i e_i$ with $a_i \in \br$. Let $\{a\} := a - \lfloor a \rfloor$ denote the fractional part of $a$. Let $\lambda = \sum_{i=1}^{r}\{a_i\}$. Hence, we can write $\sigma = \lambda + \sum_{i=1}^{r}\lfloor a_i \rfloor e_i.$ Hence, $\sigma$ lies in the $\bz_{\geq 0}$ span of $\lambda, e_1,\ldots, e_r$, which all have first coordinate less than $\sum_{i=1}^{r}\delta_i$. Therefore, $\Sigma$ is generated by elements whose first coordinate is less than $\sum_{i=1}^{r}\delta_i$.
\end{proof}

\begin{lem}
\label{lem:composite-map}
Let $D = \sum_{i=1}^{n}\alpha_i D_i$ where $D_i = V(f_i)$. Suppose we have a surjection $\phi: \bk[x_1,x_2,\ldots, x_r] \rightarrow R_D, x_i \mapsto p_i(f_1, \ldots, f_n),$ where $p_i$ is a monomial in $f_1,\ldots, f_n$. Then, define
\begin{align*}
	\Sigma = \langle u^d y_1^{c_1} \cdots y_n^{c_n} : c_i \geq -d \alpha_i, \sum_{i=1}^{n}a_ic_i = \sum_{i=1}^{} b_i c_i. \rangle 
\end{align*}
In this case, we have a factorization of $\phi$ given by maps $\chi, \psi$ defined by
\todo{Make this have mapsto on bottom}
$$\begin{tikzcd}
\bk[x_1,\ldots, x_r] \ar {r}{\chi} & \bk[\Sigma] \ar {r}{\psi} & R_D \\
x_i \ar r & u^{d_i}y_1^{c_{i1}} \cdots y_n^{c_{in}} \ar r & u^{d_i}f_1^{c_{i1}} \cdots f_n^{c_{in}}
\end{tikzcd}$$
Assuming $\chi$ is surjective, the minimal degree of generation of $\ker \phi$ is at most the maximum of the minimal degree of generation of $\ker \chi$ and the minimal degree of generation of $\ker \psi$.
\end{lem}
\begin{proof}
Note that we have an exact sequence
$$\begin{tikzcd}
0 \ar r & \ker \chi \ar r & \ker \phi \ar r & \ker \psi \ar r & 0
\end{tikzcd}.$$
To see this, note that the left map is an inclusion because $\phi = \chi \circ \psi$. Next, the right map, given by $\chi|_{\ker \phi}:\ker \phi \rightarrow \ker \psi$ is a surjection, again because $\phi = \chi \circ \psi$. The composition of the two maps is $0$ by definition of the kernel of $\chi$. Finally, if $a \in \ker \phi$ and $a$ lies in the kernel of the right map, then by definition $\chi(a) = 0$.
\end{proof}

\begin{lem}
\label{lem:bound-ker-chi}
Retaining the notation of Lemma ~\ref{lem:composite-map}, if $\Sigma$ has extremal rays $e_1,\ldots, e_r$ in degrees $d_1, \ldots, d_r$ then $\ker \chi$ has a system of generators in degrees less than $\sum_{i=1}^{r}d_r-1$.
\end{lem}
\begin{proof}
Since $e_1, \ldots, e_r$ are extremal rays, by Proposition ~\ref{prop:cone-generation}, every element $\sigma \in \Sigma$ can be written in a canonical form $\lambda + \sum_{i=1}^{r}e_i$. Now, letting $\lambda_0 := 0,\lambda_1, \ldots, \lambda_m$ be all elements of $\Sigma$ so that we can write $\lambda_i = \sum_{i=1}^{r}a_i e_i$ with $0 \leq a_i < 1$ for all $i$. Then, for any $1 \leq i \leq j \leq m,$ we can write $\lambda_i + \lambda_j$ in the above canonical form, yielding a relation in degree at most $\deg \lambda_i + \deg \lambda_j \leq 2 \cdot \left( \sum_{i=1}^{r}d_r -1 \right).$ Furthermore, these relations generate all relations, as one can apply a sequence of these relations to put any $\sigma \in \Sigma$ into canonical form.
\end{proof}

@@

\todo{possibly change this to only have two equations}
\begin{prop}
\label{prop:pm-span-and-basis}
The functions $u^d \cdot \prod_{i=1}^n f_i^{c_i}$ so that 
\begin{align}
\label{align:pm-span}
\sum_{i=0}^{n} c_i \cdot d_i \cdot f_i = 0 & \text{ and } &c_i \geq \lfloor \alpha_i d\rfloor	
\end{align}
form a spanning set for $H^0(\bp^m, dD)$ over $k$. Furthermore, functions 
of the form $u^d \cdot \prod_{i=1}^n f_i^{c_i} $ such that
\begin{align}
\label{align:pm-basis}
c_i = -\lfloor d\alpha_i \rfloor & \text{ for } & i > m
\end{align}
form a basis for $H^0(\bp^m, dD)$ over $k$.
\end{prop}
\begin{proof}
By definition of $H^0(\bp^m,dD)$, functions satisfying conditions 
~\eqref{align:pm-span} lie in $H^0(\bp^m,dD)$. To complete the proof, it suffices to check functions satisfying conditions ~\eqref{align:pm-basis} form a basis of $H^0(\bp^m,dD)$. Note that there $\binom{m+ \lfloor dD \rfloor }{m}$ functions satisfying condition ~\eqref{align:pm-basis}. However we know $h^0(\bp^m,dD) = \binom{m+ \lfloor dD \rfloor }{m},$ so it suffices to show that those functions satisfying condition ~\eqref{align:pm-basis} are independent. This follows from the assumption that $f_0,\ldots, f_n$ form a basis of degree 1 rational functions, and so degree $\lfloor d \deg D \rfloor $ monomials in $f_0,\ldots, f_n$ form a basis of degree $\lfloor d \deg D \rfloor $ rational functions.
\end{proof}
\todo{the following bound is the generation bound beginning from evan's paper}

@@

\begin{rem}
Remark about how we can assume divisor only consists of hyperplanes
by taking the Veronese embedding.
\todo{Robin: do this}
\end{rem}
